## Reproducible Research

### Peer Assessment 1
=====================

### *István Ilyés*


```{r setoptions, echo=FALSE}
opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

 *1. Loading and preprocessing the data*

```{r loading}
sourceurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
zipDestFile <- "C:\\S\\Coursera\\Reproducable research\\PA1\\activity.zip"
download.file(sourceurl, zipDestFile)
unzip(zipDestFile, overwrite = TRUE)
activity <- read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
str(activity)
```


 *2. What is mean total number of steps taken per day?*

First Make a histogram of the total number of steps taken each day.

```{r stepsday, results='hide'}
step_day  <- tapply(activity$steps, activity$date, sum)
```

```{r hist1, fig.height=4, fig.width=4}
hist(step_day, main ="Frequency of steps per day", xlab="Steps per day")
```

Then let's see the mean and the median:      
a, mean     

```{r mean}
mean_day  <-mean   (step_day, na.rm=TRUE)
mean_day
```

b, median
```{r median}
median_day<-median (step_day, na.rm=TRUE)
median_day
```

*3. What is the average daily activity pattern?*  

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

```{r prep, results='hide'}
activity$step_interval  <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
library(ggplot2)
```
```{r plot1, fig.height=4, fig.width=4}
qplot(interval,step_interval,data=activity, geom="line")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max}
activity$max_step <- max(activity$step_interval, na.rm=FALSE)
activity$interval[activity$step_interval==activity$max_step]
```

*4. Imputing missing values*

Calculating and reporting the total number of missing values in the dataset: 
```{r missing1}
colSums(is.na(activity))
```


Filling in all of the missing values in a new dataset. Strategy for imputing: using the mean per intervals

```{r impute}
activity2 <- activity
activity2$steps[is.na(activity2$steps)] <- tapply(activity2$steps, 
                                                  activity2$interval, mean, na.rm=TRUE)
```

   
Making a histogram of the total number of steps taken each day and calculating and report the mean and median total number of steps taken per day.

```{r hist2, fig.height=4, fig.width=4}
step_day2  <- tapply(activity2$steps, activity2$date, sum)
hist(step_day2, main ="Frequency of steps per day, without missing", xlab="Steps per day")

```
```{r descr2}
mean_day2  <-mean   (step_day2)
median_day2<-median (step_day2)
mean_day2
median_day2
```

The median is higher with 1 compared to the dataset where there were missing values.The histogram shows that the middle bracket of total daily steps has increased in the new dataset (between 10000-15000 steps)


*5. Are there differences in activity patterns between weekdays and weekends?*


New factor variable with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day:


```{r factor}
Sys.setlocale("LC_TIME","English")
activity2$day <- weekdays(as.Date(activity2$date))
day <- activity2$day
activity2$week_stat = ifelse(day=="Monday"| day=="Tuesday"| 
                   day=="Wednesday"|day=="Thursday"|day=="Friday",
                   "weekday", "weekend")
activity2$week = factor(activity2$week_stat,
                        labels = c("weekdays", "weekends"))
```

Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):

```{r panelplot}
library(lattice)
xyplot(activity2$step_interval ~ activity2$interval | activity2$week,
       type="l", xlab="Interval", ylab="Number of steps")
```

